\section{Introduction}
\label{sec:intro}

The 2020s have produced an unusual confluence of demands on userspace
runtimes. Confidential-compute platforms expect every cross-component
call to be brokered by a runtime with a defensible authorisation
story~\cite{tdx2024,sevsnp,arm-cca}. AI agents that drive web browsers,
shells, and developer toolchains in the field today route every
tool-call through some flavour of \texttt{bubblewrap},
container, or microVM~\cite{anthropic-cu, openai-agents, devin}, none
of which expose a capability layer above the syscall boundary.
NIST AI~600-1 and the EU AI Act both push toward
\textit{verifiable mediation} of agent actions~\cite{nist-ai-600-1,
euaiact15}, an explicit demand that runtime authorisation be
both fast and unforgeable.

The runtimes that currently come closest divide along two axes (\Cref{tab:landscape}):
\textbf{kernel-resident vs.\ userspace}, and \textbf{capability-mediated vs.\
not}. seL4~\cite{klein2009sel4} owns the kernel-capability cell; it offers
the strongest assurances but requires a clean-slate kernel and is unsuitable
for layering on Linux. gVisor~\cite{gvisor} and Firecracker~\cite{firecracker}
own the userspace-mediation cell, but neither has a capability model;
they enforce isolation through syscall interception or microVM
separation. CAP-VMs~\cite{capvms} sit in the userspace-capability cell
but require CHERI/Morello hardware. The cell that is empty---and that
the agent-runtime requirement is pushing into existence---is
\textit{userspace, capability-mediated, no special hardware, on stock
Linux}.

\sysnamefull is our entry into that cell. It is a C++ runtime that
gates every zero-copy memfd IPC channel on stock Linux through
unforgeable 128-bit \captokens with epoch-based cascading revocation,
achieving sub-microsecond validated message delivery without kernel
modifications.

\para{The empirical question}
A capability gate that costs hundreds of nanoseconds per message turns
"every message validates" into a non-starter; the runtime would lose
its IPC argument before the security one. The question this paper
answers is therefore not "can we do it?" but "can the gate sit on
the IPC fastpath without losing the latency floor that motivates
zero-copy IPC in the first place?"

We answer this with a head-to-head measurement against five reference
transports---pipe(2), socketpair, io_uring, Aeron, eRPC---in the same
TSC domain on the same machine.

\para{Why this is non-obvious}
Two design choices are easy to get wrong. First, the natural
implementation of capability validation is a hash-table lookup, which
costs at least one indirection plus a cache miss; on a sub-microsecond
budget that is most of the budget gone. Second, the natural threat
model includes \textit{forged capability identifiers}, which closes off
the obvious optimisation of trusting the token to encode its own pool
location.

Our central trick is to trust the token \emph{partially}: the high
32\,bits of the 128-bit identifier carry the pool slot index, while the
remaining 96\,bits are RDRAND-sourced randomness. The slot index is
treated as a hint, not a credential---a constant-time compare on the
full 128 bits still has to match before the token is accepted. The
forgery space stays at $2^{96}$, but the lookup is one shift plus one
load.

\para{Contributions}
\begin{enumerate}
\itemsep0pt
\item The first software-only Linux-userspace realisation of a Redell /
EROS / Cornucopia\,Reloaded\,\cite{redell74,shapiro1999eros,
cornucopia-reloaded}-class capability revocation regime, gating a
wait-free MPSC IPC fastpath in shared mmap memory
(\Cref{sec:design},~\Cref{sec:impl}).
\item The slot-indexed validate path that makes the gate affordable
on the IPC fastpath without losing the $2^{96}$ forgery space
(\Cref{sec:design:fastpath}).
\item A reproducible, same-TSC-domain head-to-head against pipe(2),
socketpair, io_uring, Aeron, and eRPC, with throughput, MPSC scaling,
revocation-latency, and microarchitectural perf-counter measurements,
all driven by a single shell command (\Cref{sec:eval}).
\item A property-based oracle test, a CBMC bounded-model proof of
capability monotonicity, and an MPSC-contention revocation test that
together back every numerical claim in the paper
(\Cref{sec:eval:correctness}).
\end{enumerate}

\Cref{sec:eval} reports per-message gate cost of \gateOverheadP99 ns at
p99 across payload sizes 64 B--16 KiB, keeping the gated path within
\aeronGapPercent\% of Aeron's RTT floor. The slot-indexed validate
itself runs in \validateP99 ns at p99 \textit{regardless of pool
occupancy} (\Cref{fig:figure-2}), confirming the $O(1)$ claim.

\para{What we explicitly do not claim}
$O(1)$ revocation is not an algorithmic novelty---Redell~\cite{redell74}
and EROS~\cite{shapiro1999eros} established the pattern, and
Cornucopia\,Reloaded~\cite{cornucopia-reloaded} is a recent peer.
Our contribution is the \textit{first software-only realisation on
stock Linux}, with measured numbers under contention. memfd zero-copy
IPC is similarly not novel as a substrate---Aeron and io_uring share
the design template; we reuse it and gate it.

\para{Roadmap}
\Cref{sec:threat} states the threat model. \Cref{sec:bg} compares
against capability OSes and modern IPC primitives. \Cref{sec:design}
develops the slot-indexed gate. \Cref{sec:impl} reports the C++
implementation, the M-21 NASM constant-time primitives, and the CBMC
proof. \Cref{sec:eval} presents the evaluation. \Cref{sec:disc}
discusses limitations including the C++ language stance.
\Cref{sec:rw} surveys related work in detail. \Cref{sec:conclusion}
concludes.
