\section{Threat model}
\label{sec:threat}

\sysname operates above the Linux kernel as a userspace runtime
shared by mutually-distrusting components within a single host. We
identify the trust boundary explicitly because Paper 1's claims live
or die by where it is drawn.

\para{Trusted}
\begin{itemize}
\itemsep0pt
\item The Linux kernel and its memory-protection guarantees. We rely
on \texttt{mmap(MAP\_SHARED)} backed by \texttt{memfd\_create}, on
\texttt{seccomp-BPF} for syscall filtering, and on
\texttt{CLONE\_NEWUSER}/\texttt{CLONE\_NEWNS} for namespace isolation.
\item The \sysname runtime's own pages, including the
\texttt{CapabilityManager} pool of token slots, the
\texttt{ChannelControlBlock}s of any IPC channels that have been
established, and the M-21 constant-time NASM primitives.
\item RDRAND as a uniformly-random source. We use 96 bits of RDRAND per
token; a compromised RDRAND would shrink the forgery space.
\end{itemize}

\para{Untrusted}
\begin{itemize}
\itemsep0pt
\item Any process holding a \captoken. The bits of the token may be
arbitrary---we assume the holder may forge, mutate, replay, or simply
guess a token's 128-bit identifier.
\item Any byte that crosses an IPC channel. The runtime must not be
induced to take any state-altering action by parsing untrusted bytes;
in particular, the IPC framing itself (header parsing, sequence
numbers) must be safe under arbitrary input.
\item The peer endpoint of a channel. A capability-bearing producer
may misbehave (publish wrong-shaped messages, withhold messages,
flood). The capability gate at the consumer protects the consumer
from the producer; the producer's gate protects \sysname from the
producer attempting to publish past its revocation.
\end{itemize}

\para{In scope: forgery, replay, use-after-revoke}
The capability check must reject:
\textit{(a)} a token whose 128-bit identifier is not currently active
in the pool (forgery);
\textit{(b)} a token whose 128-bit identifier was once active but has
since been revoked, including via cascading revocation of an ancestor
(use-after-revoke);
\textit{(c)} a token that bears the correct identifier but whose
permission bitmask does not include the requested operation (privilege
escalation).

The slot-indexed validate path (\Cref{sec:design:fastpath}) keeps the
forgery space at $2^{96}$: the slot index in the high 32 bits is
\textit{treated as a hint}, not as a credential, because the
constant-time compare on the full 128 bits still has to match.

\para{In scope: side-channels in the gate}
The capability comparison must not leak which byte differs first, and
the runtime must not take observably different paths based on a
token's contents past the binary accept/reject decision. We use the
M-21 \texttt{asm\_ct\_compare} primitive for the 128-bit identifier
match; it is constant-time and barrier-balanced. The slot index
\textit{is} consulted before the constant-time compare, but only to
choose a pool offset---an attacker who controls the slot index cannot
induce different cache behaviour across pool sizes (see
\Cref{sec:eval:perfctrs}).

\para{Out of scope: rowhammer-class faults, full kernel TCB bypass}
We do not defend against rowhammer-induced corruption of the token
pool, against speculative execution that pierces the kernel TCB
(those threats are HongMeng's territory~\cite{hongmeng}), or against
an attacker who can already write to \sysname's process memory.
\sysname assumes Linux's process isolation works.

\para{Out of scope: cross-host capability delegation}
Paper 1 is single-host. Cross-host delegation, attestation handshakes,
and remote-capability transfer are the subject of follow-up work
(M-23 ConfCompute, in preparation~\cite{astra-confcompute}).

\para{Out of scope: hardware capability enforcement}
We do not assume CHERI, Morello, MTE, or LAM. CAP-VMs~\cite{capvms}
operates in the same conceptual layer but requires CHERI hardware;
\sysname's contribution is making the same conceptual layer available
on commodity x86-64 / aarch64.
