\section{Background and related work}
\label{sec:bg}

We summarise prior work along the two axes that frame the empty cell
\sysname enters: \textbf{capability mediation} and \textbf{userspace
IPC}. Extended discussion is deferred to \Cref{sec:rw}.

\subsection{Capability systems}
\label{sec:bg:cap}

The capability discipline goes back to Dennis~\&\ Van~Horn~\cite{dennis1966}
and was made cryptographically respectable by Redell~\cite{redell74},
whose epoch-versioned credential pattern is the algorithmic ancestor
of every revocable-capability system since. KeyKOS~\cite{keykos} and
EROS~\cite{shapiro1999eros} demonstrated that the pattern works at OS
scale; CapROS~\cite{capros} carried it forward.
seL4~\cite{klein2009sel4} closed the loop with full functional
correctness verification of a capability microkernel.

In modern systems, two recent works are the immediate peer set.
\textbf{Cornucopia\,Reloaded}~\cite{cornucopia-reloaded} (ASPLOS '24)
generalises the epoch-revocation pattern to a CHERI hardware
substrate and demonstrates O(1) bulk revocation; \sysname's revocation
algorithm \emph{is} Cornucopia\,Reloaded's, ported to software and
held under the Linux process boundary. We cite this prominently: the
algorithmic novelty is theirs; our contribution is the realisation
without CHERI.
\textbf{HongMeng}~\cite{hongmeng} (OSDI '24) is the closest competitor on
the (Linux-compatible, capability-based, IPC-fastpath) triple. It is
a kernel below Linux, not a runtime above it, and its authors describe
retreating from "pure" capability chains to "address tokens" because
chain-revocation cost was hurting common-case performance. \sysname
sits on the other side of the kernel boundary; \Cref{sec:disc:hongmeng}
discusses the consequences.

\textbf{CAP-VMs}~\cite{capvms} (OSDI '22) is the closest peer on
capability + IPC, but requires CHERI/Morello hardware. \sysname's
positioning is "the same conceptual layer, on stock Linux".

\subsection{Userspace IPC primitives}
\label{sec:bg:ipc}

\textbf{Aeron}~\cite{aeron-real-logic} ships a
single-producer-single-consumer IPC over an mmap'd ring buffer with
~250\,ns published RTT for 100-byte payloads on commodity Xeon. It is
the latency floor we measure against. ByteDance's
shmipc~\cite{bytedance-shmipc} is production-deployed and
architecturally similar.

\textbf{io\_uring}~\cite{iouring} is the prototype of "mmap'd SQ/CQ
ring buffer plus futex-style notification" in the Linux kernel
itself; \sysname's RingBuffer reuses the design template at a higher
layer (the kernel doesn't see our messages; we don't pay
\texttt{enter}-time syscall cost).

\textbf{eRPC}~\cite{erpc} (NSDI '19) is the published peer-reviewed
RPC number with the lowest latency we are aware of (~2.3\,$\mu$s for a
single-hop request/response). It is an RPC layer rather than a raw
transport; we cite it for the network-RPC reference point but do not
fight on raw IPC RTT.

\textbf{LITESHIELD}~\cite{liteshield} (USENIX ATC '25) is the
architectural neighbour: a userspace $\mu$kernel-style runtime on
Linux with shared-memory IPC and a ~22-syscall host interface. It has
no capability layer. \sysname's framing is "the capability layer over
that architectural shape"; we cite LITESHIELD prominently because (a)
the architectural overlap is real, and (b) ATC '25 is the venue we
are submitting to.

\subsection{The empty cell}

\Cref{tab:landscape} summarises the cells. The intersection of
\textit{(Linux compatible)} $\cap$ \textit{(no CHERI)} $\cap$
\textit{(capability-mediated IPC)} $\cap$ \textit{(measured against
Aeron-class numbers)} is empty in the current literature. \sysname
fills it.

\begin{table}[t]
\centering\small
\begin{tabular}{lccccc}
\toprule
System            & Linux & no CHERI & Caps & IPC fastpath & Userspace \\
\midrule
seL4              & no  & yes & yes & yes & no  \\
HongMeng          & yes & yes & weak & yes & no  \\
gVisor            & yes & yes & no  & no  & yes \\
Firecracker       & yes & yes & no  & no  & yes \\
CAP-VMs           & yes & no  & yes & yes & yes \\
LITESHIELD        & yes & yes & no  & yes & yes \\
Aeron             & yes & yes & no  & yes & yes \\
io\_uring         & yes & yes & no  & yes & no  \\[2pt]
\textbf{\sysname} & \textbf{yes} & \textbf{yes} & \textbf{yes} & \textbf{yes} & \textbf{yes} \\
\bottomrule
\end{tabular}
\caption{The cell \sysname enters. "weak" caps in HongMeng marks the
authors' own retreat to address tokens.}
\label{tab:landscape}
\end{table}
