\section{Implementation}
\label{sec:impl}

\sysname is implemented in C++23 using \texttt{cmake} as the build
driver. The codebase is approximately 7,400 lines of source code
across the three subsystems we describe in this section: the
capability manager (\Mone), the IPC ring (\Mthree), and the M-21
assembly primitives.

\subsection{Capability manager (\Mone)}
\label{sec:impl:cap}

\Mone is a single C++ class, \texttt{CapabilityManager}, with a
fixed-capacity slot pool sized at construction time
(\texttt{DEFAULT\_MAX\_TOKENS}\,=\,4096). The pool is a
heap-allocated flat array of \texttt{TokenSlot} structures. There is
no dynamic allocation in the validate or revoke paths.

\para{Concurrency}
\texttt{validate()} is lock-free: it loads the global
\texttt{m\_uCurrentEpoch} with acquire ordering, performs an O(1)
slot lookup, and returns. \texttt{create()}, \texttt{derive()}, and
\texttt{revoke()} serialise through a custom spinlock
(\texttt{m\_bSpinlock}) which guards the slot-allocation, descendant
search, and bookkeeping fields. The lock release on these paths
performs a release-store of the spinlock byte, which establishes the
synchronizes-with edge any subsequent \texttt{validate()} on another
core needs to see the new state.

\para{Token-id minting}
\texttt{generateTokenId(slot, out)} fills the 128-bit identifier from
RDRAND (via the M-21 \texttt{asm\_rdrand64} wrapper) and packs the
slot index into the high 32 bits of \texttt{out[0]}. Both 64-bit
halves are checked for RDRAND failure; we fall back to an
epoch-XOR-magic-constant value with a logged warning if RDRAND fails
(this should not happen on supported microarchitectures but is
required for graceful degradation).

\subsection{Ring buffer (\Mthree)}
\label{sec:impl:ring}

\Mthree is a single-producer, single-consumer ring buffer with an
MPSC-capable claim path. The ring data and three-cache-line control
block live in a \texttt{memfd\_create} region mapped
\texttt{MAP\_SHARED} into both endpoints. The control block layout
(\Cref{fig:cb}) places the claim/commit indices and the read index
on separate cache lines to avoid false sharing.

\begin{figure}[t]
\centering\small
\begin{verbatim}
+--------------------------------+ line 0
| atomic<u64> write_index        |
| atomic<u64> write_claim_index  |
+--------------------------------+ line 1
| atomic<u64> read_index         |
+--------------------------------+ line 2
| u32 channel_id, ring_bytes ... |
+--------------------------------+ 192B total
\end{verbatim}
\caption{\texttt{ChannelControlBlock} layout. The static\_assert that
the struct is exactly 3$\times$64\,B is enforced at compile time.}
\label{fig:cb}
\end{figure}

\para{Two-phase claim/commit}
For payloads $\le 256$\,B the producer claims its byte range with a
single \texttt{fetch\_add} on \texttt{write\_claim\_index}
(wait-free); for larger payloads it CASes (lock-free with bounded
retries). After writing the data, the producer advances
\texttt{write\_index} with a CAS that spins on the prior committed
value, ensuring readers observe a contiguous ordered byte stream
regardless of how many producers raced.

\para{The capability gate}
\texttt{writeGated(\&\_gate, data, len)} calls
\texttt{gate.m\_pManager->validate(gate.m\_token, IPC\_SEND)} and
returns \texttt{PERMISSION\_DENIED} on failure, leaving
\texttt{write\_claim\_index} untouched. On success it forwards to
\texttt{write(data, len)}. The gate sits in front of the claim, so a
revoked token cannot consume claim space.

\para{The blocking variant}
\texttt{readWaitGated()} parks on \texttt{write\_index.wait()} (which
maps to a futex on Linux) and re-validates after each wake. The
extra validate is the cost of preserving correctness across
revocations that fire while the reader is asleep.

\subsection{M-21 constant-time primitives}
\label{sec:impl:asm}

The \Mtwentyone module ships seven assembly-language routines hand-written in
NASM following the System V AMD64 ABI:
\texttt{asm\_ct\_compare} (constant-time 16-byte compare),
\texttt{asm\_secure\_wipe} (memset that the optimiser cannot remove),
\texttt{asm\_rdrand64} (RDRAND wrapper),
\texttt{asm\_lfence}, \texttt{asm\_cache\_flush},
\texttt{asm\_spectre\_barrier}, \texttt{asm\_stack\_guard}.

These are not contributions of the paper---libsodium and BoringSSL
ship equivalents---but the constant-time discipline matters because
we use \texttt{asm\_ct\_compare} on the IPC fastpath. We mention this
in Implementation rather than Design because the design is
agnostic to the comparator implementation.

\subsection{CBMC monotonicity proof}
\label{sec:impl:cbmc}

We model the permission arithmetic of \texttt{CapabilityManager}
in a small C-flat translation
(\texttt{formal/cbmc/capability\_monotonicity.c}, 280 LOC) and prove
three properties using CBMC~\cite{cbmc}:

\begin{description}
\itemsep0pt
\item[Subset-on-success] If \texttt{derive(parent, child\_perms)}
returns success, then \texttt{(child\_perms \& \textasciitilde
parent\_perms) == 0}.
\item[Rejection-bijection] If \texttt{derive} returns failure, then
\texttt{(child\_perms \& \textasciitilde parent\_perms) != 0} (and
vice versa).
\item[No-recovery-through-chain] For any derivation chain
$t_0 \to t_1 \to \dots \to t_n$, the permission set of $t_n$ is a
subset of the permission set of $t_0$.
\end{description}

CBMC reports 0 of 37 checks failing on the model in
\texttt{formal/cbmc/Makefile}'s \texttt{verify} target. We frame this
as \textit{spec correspondence on the one property the runtime must
not get wrong}, not as a substitute for full functional verification
in the seL4 sense~\cite{klein2009sel4}.

\subsection{Reproducibility}
\label{sec:impl:repro}

The artefact (source, baselines, sweep script, plotting) is published
at \ifanon\textcolor{red}{[anonymous URL during review]}\else
\url{https://github.com/KernelArch-Lab/Astra}\fi.
A single shell command, \texttt{scripts/run\_paper1\_sweep.sh},
builds every harness, runs each in the same TSC domain on the same
machine, concatenates the CSV rows into
\texttt{artefact/paper1\_figure\_1.csv}, and produces Figures 1 and
2 plus Table 1 via \texttt{plot\_paper1\_figure\{,\_2\}.py}.
\Cref{tab:artefact} maps each figure and claim to its source.

\begin{table}[t]
\centering\small
\begin{tabular}{lll}
\toprule
Figure / claim         & Source script         & Output \\
\midrule
Fig 1 latency          & \texttt{run\_paper1\_sweep.sh} \tiny\texttt{+plot 1} & \texttt{paper1\_figure\_1.pdf} \\
Fig 2 O(1) validate    & \texttt{bench\_pool\_scaling}    \tiny\texttt{+plot 2} & \texttt{paper1\_figure\_2.pdf} \\
Tbl 1 256\,B comparison & \texttt{plot\_paper1\_figure.py}                       & \texttt{paper1\_table\_1.tex}  \\
Throughput             & \texttt{bench\_throughput}                              & CSV                          \\
MPSC scaling           & \texttt{bench\_throughput\_mpsc}                        & CSV                          \\
Revoke window          & \texttt{bench\_revocation\_latency}                     & CSV                          \\
Cycles / cache misses  & \texttt{bench\_perfcounters}                            & CSV                          \\
Monotonicity proof     & \texttt{make -C formal/cbmc verify}                     & CBMC report                  \\
Concurrency stress     & \texttt{test\_capability\_concurrency}                  & ctest pass                   \\
Survivability          & \texttt{test\_capability\_survivability}                & ctest pass                   \\
Gate concurrency       & \texttt{test\_sprint5\_gate\_concurrency}               & ctest pass                   \\
Property test          & \texttt{test\_capability\_property}                     & ctest pass                   \\
\bottomrule
\end{tabular}
\caption{Artefact map. Every numerical claim and every functional
correctness claim in this paper has a CI-runnable script.}
\label{tab:artefact}
\end{table}
