\section{Discussion and limitations}
\label{sec:disc}

\subsection{The C++ stance}
\label{sec:disc:cpp}

\sysname is C++ rather than Rust. We discuss the choice openly because
ATC reviewers will. The pitch is not "C++ is fine"---we have no
illusions about its memory-safety story---but rather "capability
discipline can be retrofitted onto an unsafe-language ecosystem in
ways the security literature underrepresents". \sysname demonstrates
that a 4096-token capability layer with constant-time
mediation can be added on top of a C++ runtime without the
language being a fundamental obstacle. This is a different bet from
Tock~\cite{tock}, Theseus~\cite{theseus}, and Redox, all of which take
the position that the language matters and choose Rust.

The bet is testable: the CBMC monotonicity proof, the property test
oracle, the libFuzzer harness, and the constant-time NASM primitives
together cover the surface area where C++'s unsafety is most likely
to manifest as a security bug. We disclose the residual risk---a
memory-corruption attack inside the runtime's own pages would
compromise every token---and frame Paper 1's contribution as
\textit{the upper-half capability story}, not the lower-half
safety story.

A subset port of the hot path to Rust is on our roadmap conditional
on Paper 1 reviewers flagging the language choice as the primary
weakness.

\subsection{The Aeron gap}
\label{sec:disc:aeron}

Aeron's published 250\,ns RTT is the tightest single-host IPC number
in production. Our \aeronGapPercent\% gap is the headline number
we want to defend (down) and the headline number reviewers will
attack (up). The gap decomposes as follows:

\begin{itemize}
\itemsep0pt
\item Two \texttt{validate()} calls per round trip ($\sim$\gateOverheadP99\,ns
total at p99). This is irreducible without either dropping security
or moving to a less constant-time comparator.
\item C++23 range-checked vector and \texttt{Result<T>} unwrap on the
hot path. Aeron's published number uses a slimmer, more C-like API.
A future port of the hot path to a no-std-library style would close
some of this gap; we have not done it.
\item Memory-fence policy. Aeron's ring uses release-acquire on
write\_index only; we use release-acquire on both write\_index and
write\_claim\_index. This is a correctness requirement of the
two-phase MPSC claim/commit but does cost cycles relative to the
SPSC-only Aeron arrangement.
\end{itemize}

We do not expect to beat Aeron on raw RTT---the design template is
shared---but we expect the gap to shrink with a hot-path API
streamline.

\subsection{HongMeng-overlap framing}
\label{sec:disc:hongmeng}

HongMeng (OSDI '24) is the closest competitor on the
\textit{capability + Linux + IPC fastpath} triple. The honest framing
is that HongMeng owns the kernel-level cell, while \sysname owns
the userspace-runtime cell. HongMeng's authors describe retreating
from chain-revocable capabilities to "address tokens" because the
revocation cost was hurting performance; we read this as evidence
that the (in-kernel, full-revocation) point in the design space is
expensive, not as evidence that revocation itself is unaffordable.
\sysname's userspace position pays a different cost (the slot lookup
on every message) but recovers full chain revocation at $O(1)$.

\subsection{Single-machine scope}
\label{sec:disc:scope}

Paper 1 is single-host. The cross-host story---attestation-backed
capability delegation, remote token transfer, capability tokens as
attestation artefacts in TEEs---is the subject of follow-up work
(M-23 ConfCompute) which we deliberately exclude here to keep the
contribution scoped. Reviewers who want to see TEE binding will find
it in the published architecture overview~\cite{astra-confcompute}
but not in this paper's measurement.

\subsection{Macrobench}
\label{sec:disc:macro}

We do not include a NGINX / Redis / SQLite macrobench in the
evaluation. The reason is space: at 12 pages, the head-to-head against
five reference transports plus the $O(1)$ confirmation plus
revocation latency plus perf counters fills the budget. A macrobench
would shift the headline from "the gate is cheap" to "the gate is
cheap on this specific application," which is a weaker claim. The
artefact ships the harness scaffolding; we will publish the macrobench
numbers in the artefact-evaluation appendix.
