\section{Related work — extended}
\label{sec:rw}

\subsection{Capability-based operating systems}

The capability discipline traces to Dennis~\&\
Van~Horn~\cite{dennis1966}; Hydra and StarOS in the 1970s carried it
forward. Redell's thesis~\cite{redell74} introduced the
epoch-versioned credential pattern for revocation, the algorithmic
ancestor of every capability-revocation system since. KeyKOS,
EROS~\cite{shapiro1999eros}, and CapROS~\cite{capros} demonstrated
the pattern at OS scale; Coyotos refined the typing
discipline. seL4~\cite{klein2009sel4} closed the loop with full
functional-correctness verification.

Recent capability-system work splits along the question of whether
capability identifiers should be \textit{names} (KeyKOS-style) or
\textit{addresses} (CHERI-style). \sysname's tokens are names, with
the slot index serving as a deterministic address-like hint
(\Cref{sec:design:fastpath}); we get the cache behaviour of
addresses without losing the forgeability resistance of names.

\subsection{Hardware-capability systems}

CHERI~\cite{cheri-asplos15,cheri-tcg22} is the most-developed
hardware-capability platform; CAP-VMs~\cite{capvms} layers
capability-isolation on CHERI/Morello.
Cornucopia\,Reloaded~\cite{cornucopia-reloaded} is the recent peer for
the revocation pattern itself; the algorithm \sysname uses is theirs,
but our realisation is software-only on stock Linux.
PICASSO~\cite{picasso} (S\&P~'24) takes a hardware-provenance approach
to bulk invalidation that has a similar effect.

\subsection{Userspace IPC primitives}

Aeron~\cite{aeron-real-logic} is the most-published Aeron-class
single-host IPC number we are aware of. ByteDance's
shmipc~\cite{bytedance-shmipc} is production-deployed in Service
Mesh and is architecturally similar.
io\_uring~\cite{iouring} prototyped the same template (mmap'd
SQ/CQ rings with futex notification) inside the Linux kernel; our
ring buffer is an above-the-kernel reuse of that template.

eRPC~\cite{erpc} (NSDI '19) is the published peer-reviewed RPC
number with the lowest latency we are aware of (~2.3\,$\mu$s
single-hop request/response over DPDK).
LITESHIELD~\cite{liteshield} (USENIX ATC '25) is the most-recent
userspace $\mu$kernel-style runtime on Linux; the architectural
overlap is real (ours adds the capability layer).
Cap'n~Proto's RPC layer~\cite{capnproto-rpc} sits at a higher
level---marshalling and authorization on top of a transport---and is
not a fair comparison for raw IPC RTT.

\subsection{Kernel-resident capability systems on Linux}

HongMeng~\cite{hongmeng} (OSDI '24) and CertiKOS~\cite{certikos} are
the closest kernel-resident capability systems with Linux-compatible
positioning. \sysname's contribution is sitting above them in the
stack rather than competing with them in it. Our framing borrows
from HongMeng's own authors' analysis of the tradeoff between pure
capabilities and address tokens (\Cref{sec:disc:hongmeng}).

\subsection{Confidential-compute capability binding}

Recent confidential-compute work has begun to explore capability
mediation tied to attestation reports.
ScaleTrust~\cite{scaletrust} discusses TDX-bound credentials;
Conclave~\cite{conclave} discusses SEV-SNP-sealed objects.
\sysname's design intentionally leaves this to follow-up work
(M-23~ConfCompute) and stays within the Linux process boundary.

\subsection{LLM-agent sandboxing}

The Anthropic Computer Use~\cite{anthropic-cu}, OpenAI Agents
SDK~\cite{openai-agents}, and Devin~\cite{devin} stacks all sandbox
agent tool calls; none currently expose a capability layer above the
syscall boundary. NIST AI~600-1~\cite{nist-ai-600-1} and EU AI Act
Article~15~\cite{euaiact15} demand verifiable mediation. We see
\sysname as the runtime substrate this demand will pull into
existence; the AgentGuard module (M-22 in our roadmap) is in
preparation as Paper~3.

\subsection{Formal verification}

CBMC~\cite{cbmc} bounded-model checking is the technique we use for
the monotonicity proof; SPIN~\cite{spin} would have been an
alternative for the LTL fragment, but our property is a quantifier-free
boolean and CBMC is more direct. Our scope (one property, 280 LOC C
model, 37 checks) is many orders of magnitude smaller than the
seL4 effort~\cite{klein2009sel4} and we frame it as
\textit{spec-correspondence}, not as functional correctness in seL4's
sense.
